Environmental and geographic variables are effective surrogates for genetic variation in conservation planning

### 2016-08-03 code/rmarkdown/article.Rmd
+ added refs

---
title: "Environmental and geographic variables are effective surrogates for genetic variation in conservation planning"
author:
  - name: Jeffrey O. Hanson
    affiliation: a,1
  - name: Jonathan R. Rhodes
    affiliation: b
  - name: Cynthia Riginos
    affiliation: a
  - name: Hugh P. Possingham
    affiliation: a
  - name: Richard A. Fuller
    affiliation: a
address:
  - code: a
    address: School of Biological Sciences, The University of Queensland, Brisbane, QLD, Australia
  - code: b
    address: School of Geography, Planning and Environmental Management, The University of Queensland, Brisbane, QLD, Australia
corresponding_author:
  - code: 1
  - text: "To whom correspondence should be addresssed. E-mail: jeffrey.hanson@uqconnect.edu.au"
author_contributions: |
  All authors developed the study and drafted the manuscript. JOH performed all analyses.
lead_author_surname: "Hanson"
abstract: |
  Please provide an abstract of no more than 250 words in a single paragraph. Abstracts should explain to the general reader the major contributions of the article. References in the abstract must be cited in full within the abstract itself and cited in the text.
significance: |
  Protected areas are used to preserve biodiversity patterns and processes. Since the resources available for conservation are limited, protected areas need to be sited in cost-effective locations. However, despite the stated emphasis on biodiversity processes, evolutionary processes are not often used to guide reserve selection. This is due to the fact that genetic data is often too costly to obtain and process. Here we found that environmental and geographic variables are cost-effective surrogates for genetic data. These variable can be used to identify candidate reserve systems that secure the adaptive and neutral evolutionary processes for a greater number of species. This research highlights the importance of obtaining a representative sample of species when generating broad-scale prioritisations.
acknowledgements: |
  JOH is funded by an Australian Postgraduate Award (APA) scholarship. RAF has an Australian Research Council Future Fellowship. This work was supported by the Centre of Excellence for Environmental Decisions (CEED) and the Landscape Ecology and Conservation Group (LEC) at The University of Queensland.
keywords:
  - conservation
  - biodiversity
  - AFLP
  - genetic diversity
pnas_type: pnasresearcharticle
bibliography: references.bib
csl: reference-style.csl
output:
  rticles::pnas_article:
    includes:
      in_header: [preamble.tex]
---

@@

<!--  Introduction -->
Protected areas spearhead conservation efforts [@r440]. These places buffer species from anthropogenic impacts and provide places for species to persist. However, the resources available for conservation are limited. Thus protected areas need to be sited in places that achieve conservation goals for minimal cost. To ensure that the plans for protected areas (prioritisations) are cost-effective, they are often generated using decision support tools to identify near-optimal solutions. These prioritisations often contain plans for a network of multiple protected areas so they can preserve biological processes (eg. connectivity between populations). For example, \texttt{Marxan} [@r22] uses amount-based targets that specify a minimum amount of habitat to preserve for each species to identify cost-effective solutions (eg. 100 km $^2$ of land occupied by the species needs to be secured in a reserve). To promote connecitivty among multiple reserves, \texttt{Marxan} also contains the functionality to penalise fragmented solutions (using the boundary length modifier (\textit{BLM}) parameter). However, conservation planning exercises typically assume that all individuals in the same species are equivalent.

Intra-specific genetic variation is important for long-term species persistence. As a consequence, there has been increasing interest in designing prioritisations that preserve and cultivate this aspect of biodiversity [@r67; @r3; @r61]. Although the strength of natural selection is a continuous force; broadly speaking, genetic variation can be classified as either adaptive or neutral [reviewed in @r480]. Adaptive genetic variation is associated with loci that (significantly) affect fitness. Typically, such "adaptive" loci are detected due to anomalous patterns among individuals [eg. @r456; @r487; @r480; @r490], though phenotypic-genetic associations such as quantitative trait analyses or genome-wide association measures are more likely to identify functionally important genomic regions for individual species [@rXXXX; @rXXXX]. By preserving existing patterns of adaptive variation, protected areas can ensure that populations with particularly beneficial adaptations are not lost [@r3]. In contrast, neutral genetic variation is associated with loci that do not (significantly) affect fitness. Neutral variation often reflects the evolutionary history of different populations. By preserving existing patterns of neutral variation, protected areas can safeguard against the adverse effects of low genetic diversity [eg. inbreeding depression; @r67]. Thus protected area networks should aim to preserve both adaptive and neutral patterns of genetic variation for species of conservation interest [@r67; @r3]. Despite this, genetic data are not often used to inform conservation planning exercises because such data require considerable investment to obtain and analyze for a single species [@r61; but see @r5 and @r157 for developments on using genetic data in conservation planning], and most large-scale planning exercises involve hundreds of species [eg. @rXXX]. Thus prioritisations may fail to secure a sufficient proportion of the intra-specific genetic variation of species, and ultimately fail to deliver long-term conservation outcomes.

Recently, conservation scientists have proposed using surrogate data to generate prioritisations that adequately preserve intra-specific genetic variation--without needing to utilise genetic data to acheive this--but have yet to comprehensively assess whether the proposed spatial surrogates adequately represent relevant aspects of intra-specific genetic variation. Since adaptive genetic variation is ultimately driven by extrinsic selection pressures, environmental variables have been proposed to guide reserve selection as surrogates for adaptive genetic variation [eg. environmental conditions; @r247; @r2]. By preserving a more representative sample of environmental conditions across the species' ranges, these researchers aimed to generate prioritisations that would preserve more adaptive variation. However, the effectiveness of this approach remains unverified. On the other hand, neutral genetic variation arises from a reduction in gene flow between individuals. Over the last few years, the field of phylogeography has predominantly focused on describing neutral genetic diversity, and so the landscape factors that affect neutral variation are relatively well understood [@rXXXX]. Potential surrogates for neutral variation have been based on variables that predict the levels of connectivity between planning units. For instance, the isolation by distance theory [IBD; @r259] predicts that populations further apart will have less connectivity, exchange less genetic material, and in turn have less genetic variation in common. Based on this theory, geographic distance-based surrogates have been found to improve prioritisations in an island network [@r200]. However this remains untested in spatially contiguous systems where connectivity is complicated by additional factors [eg. terrain or habitat type; @r474; @r497]--the very systems commonly featured in real-world conservation planning exercises [eg. @r197]. Before conservation planners can reliably use surrogates for genetic data in conservation planning, the effectiveness of these surrogates need to be verified. Otherwise, conservation planners may waste precious resources on protecting places that provide little benefit to biodiversity.

Here we investigated the effectiveness of environmental and geographic surrogates for adaptive and neutral genetic variation in the context of conservation planning. To achieve this, we developed a novel framework for incorporating adaptive and neutral genetic data into conservation planning. First, we aim to verify if the spatial patterns in environmental and geographic variation correlate with the broad-scale spatial patterns in adaptive and neutral variation (respectively). Second, we aim to test whether environmental and geographic variables are effective surrogates for adaptive and neutral variation (respectively) when generating prioritisations. We expect that prioritisations generated using surrogate-based targets will secure more genetic variation than prioritisations generated using only amount-based targets with the same number of planning units. Third, we aim to determine if the addition of surrogate-based targets to conservation planning exercises can result in more effective prioritisations. We expect that prioritisations generated using amount- and surrogate-based targets will sufficiently preserve the genetic variation of more species than those generated using only amount-based targets. Furthermore, we aim to test this hypothesis under several of different scenarios to see whether the additional constraints of real-world planning exercises exacerbate or nullify this effect. If these surrogates can improve the intra-specific genetic variation captured in a prioritisation, then conservation planners may use these surrogates when genetic data is not available, and in turn deliver more effective protected areas.

@@

<!-- surrogate/genetic dissimilarity matrix correlations -->
Differences in the the environmental conditions and geographic position among the planning units tended to correlate with genetic differences among the individuals in the planning units (Table 1). For both surrogates and their corresponding models, the conditional $R^2$ values (environmental vs. adaptive genetic variation: mean `r round(mean(correlation.model.summary.DF$Environmental.Conditional.R2,na.rm=TRUE),2)` conditional $R^2$ `r round(sd(correlation.model.summary.DF$Environmental.Conditional.R2,na.rm=TRUE),2)` $\pm$ SD; geographic vs. neutral genetic variation: mean `r round(mean(correlation.model.summary.DF$Geographic.Conditional.R2,na.rm=TRUE),2)` conditional $R^2$ `r round(sd(correlation.model.summary.DF$Geographic.Conditional.R2,na.rm=TRUE),2)` $\pm$ SD) were generally slightly higher than the marginal $R^2$ values (environmental vs. adaptive genetic variation: mean `r round(mean(correlation.model.summary.DF$Environmental.Marginal.R2,na.rm=TRUE),2)` marginal $R^2$ `r round(sd(correlation.model.summary.DF$Environmental.Marginal.R2,na.rm=TRUE),2)` $\pm$ SD; geographic vs. neutral genetic variation: mean `r round(mean(correlation.model.summary.DF$Geographic.Marginal.R2,na.rm=TRUE),2)` marginal $R^2$ `r round(sd(correlation.model.summary.DF$Geographic.Marginal.R2,na.rm=TRUE),2)` $\pm$ SD). This result suggests that while variation in the surrogate variables correlates with genetic variation, both forms of genetic variation contain additional spatially auto-correlated variation that is not explained by these surrogates. In particular, the differences between the conditional and marginal $R^2$ values for the models examining the effectiveness of geographic variables suggest that this unexplained spatial auto-correlation may be non-linear or driven by a few neighboring planning units with genetically similar individuals. This finding is consonant with other studies that have found strong patterns of spatial auto-correlation in patterns of adaptive and neutral genetic variation [@rXXX; @rXXX; @rXXX]. Although correlations between the spatial patterns of the environmental and geographic variables and genetic variables suggest that they may be useful for genetic variation in conservation planning, these correlations alone are insufficient to demonstrate that they are actually effective in the context of conservation planning. To demonstrate their effectiveness would require examining the relationship between the proportion of surrogate variation and genetic variation secured in a given prioritisation.

<!-- surrogate/genetic held correlations -->
Prioritisations that secured a more representative sample of environmental conditions or geographic position tended to secure a greater proportion of adaptive and neutral genetic variation (Figure 2). Generally, the environmental and geographic surrogates variables were equally effective surrogates for adaptive (mean `r round(mean(env.surrogate.spp.DF$r2CU),2)` $R^2$ $\pm$ `r round(sd(env.surrogate.spp.DF$r2CU),2)` SD; Figures 2c and S56) and neutral genetic variation (mean `r round(mean(geo.surrogate.spp.DF$r2CU),2)` $R^2$ $\pm$ `r round(sd(geo.surrogate.spp.DF$r2CU),2)` SD; Figure S57). But for a few species, the proportion of geographic variation secured in a prioritisation had a diminished capacity to explain the proportion of neutral variation secured in the prioritisation (eg. _Gypsophila repens_: $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Gypsophila_repens')$Geographic.r2CU,2)`; _Luzula alpinopilosa_: $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Luzula_alpinopilosa')$Geographic.r2CU,2)`; _Gentiana nivalis_ $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Gentiana_nivalis')$Geographic.r2CU,2)`). This bimodal distribution (Figure 2d) suggests that geographic position alone may not be a suitable surrogate for neutral variation in conservation planning for all species. This result could be attributed to multiple sources. First, geographic distances will be a poor surrogate for genetic lineages in the process of sympatric speciation. Second, connectivity between areas can be affected by local topology. In particular, mountains can act as barriers to gene flow in alpine landscapes and so nearby populations may be isolated from each other. Distances that Incorporated information on local topology (eg. least-cost distances) may perform better, but generally require species-specific scaling to adequately describe connectivity. Third, the species used in this study utilise different dispersal mechanisms [wind dispersal, animal dispersal, offshoots from the mother plant; @rXXX], and some of these species may have dispersal modes that are non-linearly related to geographic distance (eg. wind dispersing plants can have high connectivity between nearby and distant areas). Overall, however, these species were the exception, and both simple surrogates performed moderately well.

@@

All the multi-species prioritisations secured a representative sample of most of the species' intra-specific adaptive genetic variation (Figures 4e and 4g). Unlike the single-species prioritisations, the multi-species prioiritisations generated with only amount-targets did not secure a representative sample of adaptive variation due to stochastic processes. Since these prioritisations were forced to secure at least `r raptr.params.LST[[MODE]]$scenario.analysis$amount.target*100` % of land occupied each species, these prioritisations were forced to secured planning units a range of environmental conditions due to the presence of habitat specialists in the data set (eg. _Cerastium uniflorum_ and _Androsace obtusifolia_). Overall--contrary to our original hypotheses--it seems that conventional reserve selection methods may secure a representative sample of species' adaptive genetic variation if a comprehensive set of species is used. This outcome provides empirical support for one of the fundamental tenants of systematic conservation planning [@rXXX].

@@

The first outlier detection method we used involved fitting multinominal-Dirichlet models implemented in \texttt{BayeScan} [version 2.1; @r456]. We adopted a similar methodology to Bothwell _et al._ [-@r455] and applied it to each of the species in the data set. Following their methodology, we first grouped conspecifics into genetic lineages to avoid falsely classifying loci as adaptive. To complete this first step, for each species, we fit the admixture and correlated alleles model implemented in \texttt{Structure} [version 2.3.4; @r461; @r462; `r structure.params.LST[[MODE]]$numruns` replicates per K; `r structure.params.LST[[MODE]]$admburnin` admixture burnin iterations; `r structure.params.LST[[MODE]]$burnin` burnin iterations; `r structure.params.LST[[MODE]]$numreps` iterations used for inference] using the number of populations for each species previously determined by Alvarez _et al._ [-@r478]. We used \texttt{ClumPP} [version 1.1.2; @r459] to combine replicate runs to group individuals into populations (greedy algorithm using the $G^{\prime}$ static; `r clumpp.params.LST[[MODE]]$repeats` iterations). We then ran $\texttt{Bayescan}$ [@r456] for each species using these clusters (1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations). Additionally, we omitted individuals from the \texttt{Bayescan} analysis if their population membership was uncertain (maximum membership probability $< `r structure.params.LST[[MODE]]$probability.threshold`$). We ran `r bayescan.params.LST[[MODE]]$reps` replicates per species using a suitable false discovery rate [$q \leq `r bayescan.params.LST[[MODE]]$fdr`$; @r466].

>>>>

---
title: "Environmental and geographic variables are effective surrogates for genetic variation in conservation planning"
author:
  - name: Jeffrey O. Hanson
    affiliation: a,1
  - name: Jonathan R. Rhodes
    affiliation: b
  - name: Cynthia Riginos
    affiliation: a
  - name: Hugh P. Possingham
    affiliation: a
  - name: Richard A. Fuller
    affiliation: a
address:
  - code: a
    address: School of Biological Sciences, The University of Queensland, Brisbane, QLD, Australia
  - code: b
    address: School of Geography, Planning and Environmental Management, The University of Queensland, Brisbane, QLD, Australia
corresponding_author:
  - code: 1
  - text: "To whom correspondence should be addressed. E-mail: jeffrey.hanson@uqconnect.edu.au"
author_contributions: |
  All authors developed the study and drafted the manuscript. JOH performed all analyses.
lead_author_surname: "Hanson"
abstract: |
  Protected areas buffer species from anthropogenic threats and provide places for processes to continue. However, evolutionary processes are not often explicitly considered when deciding where to site protected areas because genetic data require considerable resources to obtain and analyse. Here we show that environmental and geographic distance variables are effective surrogates for adaptive and neutral intra-specific variation for conservation planning, and that these surrogates can be used to deliver more effective solutions. We obtained distribution and genetic data from the international IntraBioDiv project for 27 species collected over the European Alps. Samples were collected using a gridded sampling scheme. We used these cells as planning units to generate solutions. For each species, we identified loci that were potentially under selection using outlier loci methods, and mapped their main gradients of adaptive and neutral genetic variation across the planning units. Our first analysis confirms that the the spatial patterns of adaptive and neutral genetic variation correlate with those of environmental and geographic variation (respectively). Our second analysis verifies that solutions that secure more adaptive genetic variation also tend to secure more adaptive variation, and similar patterns were found for neutral genetic variation and geographic variables. Our third analysis found that solutions generated using additional targets to explicitly secure a representative sample of the environmental and geographic variation across a species range resulted in solutions that secure more neutral genetic variation for most species. Additionally, we found that most solutions--even those generated without surrogate or genetic data--tended to secure most of the adaptive genetic variation for most species. Overall, our analysis provides a rigorous examination of two potential surrogates for genetic variation. 
significance: |
  Protected areas are used to preserve biodiversity patterns and processes. Since the resources available for conservation are limited, protected areas need to be sited in cost-effective locations. However, despite the stated emphasis on biodiversity processes, evolutionary processes are not often used to guide reserve selection. This is due to the fact that genetic data is often too costly to obtain and process. Here we found that environmental and geographic variables are cost-effective surrogates for genetic data. These variable can be used to deliver candidate reserve systems that secure the adaptive and neutral evolutionary processes for a greater number of species with minimal additional cost. This research highlights the importance of explicitly considering biodiversity processes during conservation planning exercises.
acknowledgements: |
  JOH is funded by an Australian Postgraduate Award (APA) scholarship. RAF has an Australian Research Council Future Fellowship. This work was supported by the Centre of Excellence for Environmental Decisions (CEED) and the Landscape Ecology and Conservation Group (LEC) at The University of Queensland.
keywords:
  - conservation
  - biodiversity
  - AFLP
  - genetic diversity
pnas_type: pnasresearcharticle
bibliography: references.bib
csl: reference-style.csl
output:
  rticles::pnas_article:
    includes:
      in_header: [preamble.tex]
---

@@

<!--  Introduction -->
Protected areas spearhead conservation efforts [@r440]. These places buffer species from anthropogenic impacts and provide places for species to persist. However, the resources available for conservation are limited. Thus protected areas need to be sited in places that achieve conservation goals for minimal cost. To ensure that the plans for protected areas (prioritisations) are cost-effective, they are often generated using decision support tools to identify near-optimal solutions. These prioritisations often contain plans for a network of multiple protected areas so they can preserve biological processes. For example, \texttt{Marxan} [@r22] uses amount-based targets that specify a minimum amount of habitat to preserve for each species to identify cost-effective solutions (eg. 100 km $^2$ of land occupied by the species needs to be secured in a reserve). To promote connectivity among multiple reserves, \texttt{Marxan} also contains the functionality to penalise fragmented solutions (using the boundary length modifier (\textit{BLM}) parameter). However, conservation planning exercises typically assume that all individuals in the same species are equivalent.

Intra-specific genetic variation is important for long-term species persistence. As a consequence, there has been increasing interest in designing prioritisations that preserve and cultivate this aspect of biodiversity [@r67; @r3; @r61; @r499]. Although the strength of natural selection is a continuous force; broadly speaking, genetic variation can be classified as either adaptive or neutral [reviewed in @r480]. Adaptive genetic variation is associated with loci that (significantly) affect fitness. Typically, such "adaptive" loci are detected due to anomalous patterns among individuals [eg. @r456; @r487; @r480; @r490], though phenotypic-genetic associations such as quantitative trait analyses or genome-wide association measures are more likely to identify functionally important genomic regions for individual species [eg. @r500]. By preserving existing patterns of adaptive variation, protected areas can ensure that populations with particularly beneficial adaptations are not lost [@r3; @r43]. In contrast, neutral genetic variation is associated with loci that do not (significantly) affect fitness. Neutral variation often reflects the evolutionary history of different populations. By preserving existing patterns of neutral variation, protected areas can safeguard against the adverse effects of low genetic diversity [eg. inbreeding depression; @r67] . Thus protected area networks should aim to preserve both adaptive and neutral patterns of genetic variation for species of conservation interest [@r67; @r3; @r43]. Despite this, genetic data are not often used to inform conservation planning exercises because such data require considerable investment to obtain and analyze [@r61; but see @r5 and @r157 for developments on using genetic data in conservation planning]. Thus prioritisations may fail to secure a sufficient proportion of the intra-specific genetic variation of species, and ultimately fail to deliver long-term conservation outcomes.

Recently, conservation scientists have proposed using surrogate data to generate prioritisations that adequately preserve intra-specific genetic variation--without needing to utilise genetic data to acheive this--but have yet to comprehensively assess whether the proposed spatial surrogates adequately represent relevant aspects of intra-specific genetic variation. Since adaptive genetic variation is ultimately driven by extrinsic selection pressures, environmental variables have been proposed to guide reserve selection as surrogates for adaptive genetic variation [eg. environmental conditions; @r247; @r2]. By preserving a more representative sample of environmental conditions across the species' ranges, these researchers aimed to generate prioritisations that would preserve more adaptive variation. However, the effectiveness of this approach remains unverified. On the other hand, neutral genetic variation arises from a reduction in gene flow between individuals. Over the last few years, the field of phylogeography has predominantly focused on describing neutral genetic diversity, and so the landscape factors that affect neutral variation are relatively well understood [@r501]. Potential surrogates for neutral variation have been based on variables that predict the levels of connectivity between planning units. For instance, the isolation by distance theory [IBD; @r259] predicts that populations further apart will have less connectivity, exchange less genetic material, and in turn have less genetic variation in common. Based on this theory, geographic distance-based surrogates have been found to improve prioritisations in an island network [@r200]. However this remains untested in spatially contiguous systems where connectivity is complicated by additional factors [eg. terrain or habitat type; @r474; @r497]--the very systems commonly featured in real-world conservation planning exercises [eg. @r197]. Before conservation planners can reliably use surrogates for genetic data in conservation planning, the effectiveness of these surrogates need to be verified. Otherwise, conservation planners may waste precious resources on protecting places that provide little benefit to biodiversity.

Here we investigated the effectiveness of environmental and geographic surrogates for adaptive and neutral genetic variation in the context of conservation planning. To achieve this, we developed a novel framework for incorporating adaptive and neutral genetic data into conservation planning. First, we aim to verify if the spatial patterns in environmental and geographic variation correlate with the broad-scale spatial patterns in adaptive and neutral variation (respectively). Second, we aim to test whether environmental and geographic variables are effective surrogates for adaptive and neutral variation (respectively) when generating prioritisations. Third, we aim to determine if the addition of surrogate-based targets to conservation planning exercises can result in more effective prioritisations. We expect that prioritisations generated using amount- and surrogate-based targets will sufficiently preserve the genetic variation of more species than prioritisations generated using only amount-based targets. Furthermore, we aim to test this hypothesis under several of different scenarios to see if our results are robust. If these surrogates can improve the intra-specific genetic variation captured in a prioritisation, then conservation planners may use these surrogates when genetic data is not available, and in turn deliver more effective protected areas.

@@

<!-- surrogate/genetic dissimilarity matrix correlations -->
Differences in the the environmental conditions and geographic position among the planning units tended to correlate with genetic differences among the individuals in the planning units (Table 1). For both surrogates and their corresponding models, the conditional $R^2$ values (environmental vs. adaptive genetic variation: mean `r round(mean(correlation.model.summary.DF$Environmental.Conditional.R2,na.rm=TRUE),2)` conditional $R^2$ `r round(sd(correlation.model.summary.DF$Environmental.Conditional.R2,na.rm=TRUE),2)` $\pm$ SD; geographic vs. neutral genetic variation: mean `r round(mean(correlation.model.summary.DF$Geographic.Conditional.R2,na.rm=TRUE),2)` conditional $R^2$ `r round(sd(correlation.model.summary.DF$Geographic.Conditional.R2,na.rm=TRUE),2)` $\pm$ SD) were generally slightly higher than the marginal $R^2$ values (environmental vs. adaptive genetic variation: mean `r round(mean(correlation.model.summary.DF$Environmental.Marginal.R2,na.rm=TRUE),2)` marginal $R^2$ `r round(sd(correlation.model.summary.DF$Environmental.Marginal.R2,na.rm=TRUE),2)` $\pm$ SD; geographic vs. neutral genetic variation: mean `r round(mean(correlation.model.summary.DF$Geographic.Marginal.R2,na.rm=TRUE),2)` marginal $R^2$ `r round(sd(correlation.model.summary.DF$Geographic.Marginal.R2,na.rm=TRUE),2)` $\pm$ SD). This result suggests that while variation in the surrogate variables correlates with genetic variation, both forms of genetic variation contain additional spatially auto-correlated variation that is not explained by these surrogates. In particular, the differences between the conditional and marginal $R^2$ values for the models examining the effectiveness of geographic variables suggest that this unexplained spatial auto-correlation may be non-linear or driven by a few neighboring planning units with genetically similar individuals. This finding is consonant with other studies that have found strong patterns of spatial auto-correlation in patterns of genetic variation [@r502]. Although correlations between the spatial patterns of the environmental and geographic variables and genetic variables suggest that they may be useful for genetic variation in conservation planning, these correlations alone are insufficient to demonstrate that they are actually effective in the context of conservation planning. To demonstrate their effectiveness would require examining the relationship between the proportion of surrogate variation and genetic variation secured in a given prioritisation.

<!-- surrogate/genetic held correlations -->
Prioritisations that secured a more representative sample of environmental conditions or geographic position tended to secure a greater proportion of adaptive and neutral genetic variation (Figure 2). Generally, the environmental and geographic surrogates variables were equally effective surrogates for adaptive (mean `r round(mean(env.surrogate.spp.DF$r2CU),2)` $R^2$ $\pm$ `r round(sd(env.surrogate.spp.DF$r2CU),2)` SD; Figures 2c and S56) and neutral genetic variation (mean `r round(mean(geo.surrogate.spp.DF$r2CU),2)` $R^2$ $\pm$ `r round(sd(geo.surrogate.spp.DF$r2CU),2)` SD; Figure S57). But for a few species, the proportion of geographic variation secured in a prioritisation had a diminished capacity to explain the proportion of neutral variation secured in the prioritisation (eg. _Gypsophila repens_: $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Gypsophila_repens')$Geographic.r2CU,2)`; _Luzula alpinopilosa_: $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Luzula_alpinopilosa')$Geographic.r2CU,2)`; _Gentiana nivalis_ $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Gentiana_nivalis')$Geographic.r2CU,2)`). This bimodal distribution (Figure 2d) suggests that geographic position alone may not be a suitable surrogate for neutral variation in conservation planning for all species. This result could be attributed to multiple sources. First, geographic distances will be a poor surrogate for genetic lineages in the process of sympatric speciation. Second, connectivity between areas can be affected by local topology. In particular, mountains can act as barriers to gene flow in alpine landscapes and so nearby populations may be isolated from each other. Distances that incorporate information on local topology (eg. least-cost distances) may perform better, but generally require species-specific scaling to adequately describe connectivity. Third, the species used in this study utilise different dispersal mechanisms [animal-, wind-, or gravity-based dispersal @r451], and some of these species may have dispersal modes that are non-linearly related to geographic distance (eg. wind dispersing plants can have high connectivity between nearby and distant areas). Overall, however, these species were the exception, and both simple surrogates performed moderately well.

@@

All the multi-species prioritisations secured a representative sample of most of the species' intra-specific adaptive genetic variation (Figures 4e and 4g). Unlike the single-species prioritisations, the multi-species prioiritisations generated with only amount-targets did not secure a representative sample of adaptive variation due to stochastic processes. Since these prioritisations were forced to secure at least `r raptr.params.LST[[MODE]]$scenario.analysis$amount.target*100` % of land occupied each species, these prioritisations were forced to secured planning units a range of environmental conditions due to the presence of habitat specialists in the data set (eg. _Cerastium uniflorum_ and _Androsace obtusifolia_). Overall--contrary to our original hypotheses--it seems that conventional reserve selection methods may secure a representative sample of species' adaptive genetic variation if a comprehensive set of species is used. This outcome provides empirical support for one of the fundamental tenants of systematic conservation planning.

@@

The first outlier detection method we used involved fitting multinominal-Dirichlet models implemented in \texttt{BayeScan} [version 2.1; @r456]. We adopted a similar methodology to Bothwell _et al._ [-@r455] and applied it to each of the species in the data set. Following their methodology, we first grouped conspecifics into genetic lineages to avoid falsely classifying loci as adaptive. To complete this first step, for each species, we fit the admixture and correlated alleles model implemented in \texttt{Structure} [version 2.3.4; @r461; @r462; `r structure.params.LST[[MODE]]$numruns` replicates per K; `r structure.params.LST[[MODE]]$admburnin` admixture burnin iterations; `r structure.params.LST[[MODE]]$burnin` burnin iterations; `r structure.params.LST[[MODE]]$numreps` iterations used for inference] using the number of populations for each species previously determined by Alvarez _et al._ [-@r478]. We used \texttt{ClumPP} [version 1.1.2, @r459] to combine replicate runs to group individuals into populations (greedy algorithm using the $G^{\prime}$ static; `r clumpp.params.LST[[MODE]]$repeats` iterations). We then ran $\texttt{Bayescan}$ [@r456] for each species using these clusters (1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations). Additionally, we omitted individuals from the \texttt{Bayescan} analysis if their population membership was uncertain (maximum membership probability $< `r structure.params.LST[[MODE]]$probability.threshold`$). We ran `r bayescan.params.LST[[MODE]]$reps` replicates per species using a suitable false discovery rate [$q \leq `r bayescan.params.LST[[MODE]]$fdr`$; @r466].
